### 2013-05-19 doc/dissertation/document/chapters/spn_equations/spn_equations.tex
adding to ilut section and updating ilut data

\subsection{ILUT Preconditioning}
\label{subsec:spn_ilut_preconditioning}
Incomplete lower-upper (ILU) factorizations of the linear operator can
be used a simple mechanism to form an approximate inverse as a
preconditioner. To build the factorization, the sparse upper and lower
triangular factors, $\mathbf{L}$ and $\mathbf{U}$, are computed such
that the residual matrix formed by the factorization
\citep{saad_iterative_2003}:
\begin{equation}
  \mathbf{R} = \mathbf{L} \mathbf{U} - \mathbf{A} \:,
  \label{eq:ilu_residual_matrix}
\end{equation}
has a specified sparsity sparsity pattern and element magnitude.

\subsection{Sparse Approximate Inverse Preconditioning}
\label{subsec:spn_spainv_preconditioning}

>>>>

\subsection{ILUT Preconditioning}
\label{subsec:spn_ilut_preconditioning}
Incomplete lower-upper (ILU) factorizations of the linear operator can
be used a simple mechanism to form an approximate inverse as a
preconditioner. To build the factorization, the sparse upper and lower
triangular factors, $\mathbf{L}$ and $\mathbf{U}$, are computed such
that the residual matrix formed by the factorization
\citep{saad_iterative_2003}:
\begin{equation}
  \mathbf{R} = \mathbf{L} \mathbf{U} - \mathbf{A} \:,
  \label{eq:ilu_residual_matrix}
\end{equation}
has a specified sparsity pattern and element magnitude. When a
magnitude threshold is used, elements generated in the factors below
that maginitude are dropped, resulting in ILU threshold (ILUT)
preconditioning. The sparsity pattern in this case is determined from
the input matrix to be preconditioned and the number of elements
maintained in the factor is specified by a fill level parameter. A
fill level of 1 will generate the same number of elements as the
sparsity pattern of the input matrix while a fill level of 2 will
contain twice as many elements in the factorization, resulting in a
better representation of the true LU factorization. The inverse of the
lower and upper triangular factors may then be easily inverted by
means of simple elimination to produce the preconditioner. For MCSA,
$\mathbf{L}^{-1}$ will be used on the left and $\mathbf{U}^{-1}$ on
the right to precondition the system.

For modern subspace methods, only the action to the preconditioner on
a vector is required for efficient implementations and therefore a
triangular solve can be used for this purpose. For the explicit
preconditioning scheme presented in Eq~(\ref{eq:left_right_mcsa}), the
fully inverted operator must be generated in order to build the set of
probabilities and weights required for Monte Carlo
sampling. Therefore, for a linear system of size $N$, $N$ triangular
solves will be required in order to extract the inverse matrices from
production ILU implementations. In addition, parallel implementations
typically generate lower and upper factors that are only triangular
locally, providing an easily parallelizable mechanism to generate the
action of the inverse. A consequence of this choice for parallel
scalability is a degradation of the preconditioning quality as the
size of the parallel system is increased. For serial computations, the
triangular factorization is potentially exact depending on the
parameters chosen while at thousands of parallel tasks, the global
triangular factors differ signficantly from the true factorization. As
a result, more iterations are required at higher levels of parallelism
to converge the system and will ultimately degrade overall scalability
of the system with respect to total wall time to converge.

MCSA preconditioned with ILUT was used to solve the fuel assembly
problem presented in the previous section for a 1-group $SP_1$
discretization. Unlike the Jacobi preconditioned strategy, convergence
was achieved with ILUT preconditioning. To study convergence
sensitivity to ILUT parameters, the fill level and drop tolerance were
parametrically varied with the number of iterations to converge
recorded along with the maximum number of non-zero entries observed in
all matrix rows. To provide some sparsity to the factorization, the
ILUT drop tolerance was used to drop elements in the extracted inverse
factors. For each calculation, the number of iterations required to
converge reported was for the first eigenvalue iteration with
\sn{3}{4} histories at each MCSA iteration to compute the Monte Carlo
correction.

\subsection{Sparse Approximate Inverse Preconditioning}
\label{subsec:spn_spainv_preconditioning}
Explicit formation of the preconditioner inverse is requirement of
MCSA preconditioning.
